---
title: "combine_data"
author: "Arda Ergin"
date: "2025-02-18"
output: html_document
---

## Setup
```{r}
# Install required packages if not already installed
# install.packages(c("jsonlite", "dplyr", "purrr"))

library(jsonlite)
library(dplyr)
library(purrr)
```

## Data folder selection
```{r}
# Define the folder where your JSON files are stored
folder_path <- "../data"
json_files <- list.files(path = folder_path, pattern = "\\.json$", full.names = TRUE)

# Check if this list is non-empty
# if empty, check `getwd()`,
# and set currrent working directory to the correct location, 
# the location should be where this combine_data.Rmd is located
# i.e., "......../GitHub/Comparative-judgement/analysis"
```

## Main loop
**Aim**: to create two data frames:
- One with *long-format* with all trial-level info as well as participant-level.
- One with *wide-format* that has participant-level info, but also some trial-level info that is aggregated: i.e., average reaction time, percentage of choosing left (pressing "d") for each round type, 

```{r}
# Initialize empty lists for trial-level and participant-level data
trial_rows <- list()
participant_rows <- list()

# Loop over each JSON file
for (file in json_files) {
  # Read the JSON file
  data <- jsonlite::fromJSON(file, simplifyVector = FALSE)
  
  # Extract participant-level info
  participant_id <- data$participant_id
  start_time     <- data$start_time
  end_time       <- data$end_time
  duration       <- data$duration
  
  # Extract demographics (with defaults if keys are missing)
  demographics  <- data$demographics %||% list()
  gender        <- demographics$gender %||% NA
  age           <- demographics$age %||% NA
  nationality   <- demographics$nationality %||% NA
  diet          <- demographics$diet %||% NA
  eat_frequency <- demographics$eat_frequency %||% NA
  
  # Compute overall summary measures for the participant
  trial_responses <- sapply(data$trials, function(trial) trial$response)
  reaction_times  <- sapply(data$trials, function(trial) trial$reaction_time)
  avg_reaction_time <- mean(as.numeric(reaction_times), na.rm = TRUE)
  
  # Compute percentage of missed trials (i.e., when response is "missed")
  total_trials <- length(trial_responses)
  missed_perc <- if (total_trials > 0) {
    sum(trial_responses == "missed") / total_trials * 100
  } else {
    NA
  }
  
  # Function to compute percentage of "d" responses for a given round type
  get_d_percentage <- function(round_type_val) {
    # Filter trials matching the round type
    responses <- sapply(data$trials, function(trial) {
      if (trial$round_type == round_type_val) trial$response else NA
    })
    # Remove NAs (i.e., trials that are not of this round_type)
    responses <- responses[!is.na(responses)]
    total <- length(responses)
    if(total > 0) {
      d_count <- sum(responses == "d", na.rm = TRUE)
      return(d_count / total * 100)
    } else {
      return(NA)
    }
  }
  
  # Calculate percentages for each round type
  d_perc_practice   <- get_d_percentage("practice")
  d_perc_similarity <- get_d_percentage("similarity")
  d_perc_liking     <- get_d_percentage("liking")
  
  # Create and store a participant-level row
  participant_row <- list(
    participant_id    = participant_id,
    start_time        = start_time,
    end_time          = end_time,
    duration          = duration,
    gender            = gender,
    age               = age,
    nationality       = nationality,
    diet              = diet,
    eat_frequency     = eat_frequency,
    avg_reaction_time = avg_reaction_time,
    missed_perc       = missed_perc,
    d_perc_practice   = d_perc_practice,
    d_perc_similarity = d_perc_similarity,
    d_perc_liking     = d_perc_liking
  )
  participant_rows <- append(participant_rows, list(participant_row))
  
  # Iterate over the trials to create trial-level rows
  for (trial in data$trials) {
    trial_row <- list(
      participant_id   = participant_id,
      trial_num        = trial$trial_num,
      round_type       = trial$round_type,
      left_stimulus    = trial$left_stimulus,
      right_stimulus   = trial$right_stimulus,
      comparison_order = trial$comparison_order,
      response         = trial$response,
      reaction_time    = trial$reaction_time,
      start_time       = start_time,
      end_time         = end_time,
      duration         = duration,
      gender           = gender,
      age              = age,
      nationality      = nationality,
      diet             = diet,
      eat_frequency    = eat_frequency
    )
    
    trial_rows <- append(trial_rows, list(trial_row))
  }
}

# Combine all rows into data frames
df_long <- dplyr::bind_rows(trial_rows)
df_wide <- dplyr::bind_rows(participant_rows)
```

## Convert Column Types
```{r}
# For the long-format data frame:
df_long <- df_long %>%
  mutate(
    participant_id   = as.factor(participant_id),
    trial_num        = as.numeric(trial_num),
    round_type       = factor(round_type, levels = c("practice", "similarity", "liking")),
    left_stimulus    = as.character(left_stimulus),
    right_stimulus   = as.character(right_stimulus),
    comparison_order = factor(comparison_order, 
                              levels = c(1, 2)),
    response         = factor(response, 
                              levels = c("d", "k", "missed"), 
                              labels = c("left", "right", "missed")),
    reaction_time    = as.numeric(reaction_time),
    start_time       = as.character(start_time),
    end_time         = as.character(end_time),
    duration         = as.numeric(duration),
    gender           = factor(gender, 
                               levels = c(1,2,3,4), 
                               labels = c("Male", "Female", "Non-binary", "Prefer not to say")),
    age              = as.numeric(age),
    nationality      = as.factor(nationality),
    diet             = factor(diet, 
                               levels = c(1,2,3,4,5), 
                               labels = c("Omnivore",
                                          "Vegetarian",
                                          "Vegan",
                                          "Pescatarian",
                                          "Other")),
    eat_frequency    = factor(eat_frequency, 
                               levels = c(1,2,3,4,5,6), 
                               labels = c("Never",
                                          "Once every three months",
                                          "Once a month",
                                          "Once a week",
                                          "A couple of times per week",
                                          "Every day")),
  )

# For the wide-format (participant-level) data frame:
df_wide <- df_wide %>%
  mutate(
    participant_id    = as.factor(participant_id),
    start_time        = as.character(start_time),
    end_time          = as.character(end_time),
    duration          = as.numeric(duration),
    gender            = factor(gender, 
                               levels = c(1,2,3,4), 
                               labels = c("Male", "Female", "Non-binary", "Prefer not to say")),
    age               = as.numeric(age),
    nationality       = as.factor(nationality),
    diet              = factor(diet, 
                               levels = c(1,2,3,4,5), 
                               labels = c("Omnivore",
                                          "Vegetarian",
                                          "Vegan",
                                          "Pescatarian",
                                          "Other")),
    eat_frequency     = factor(eat_frequency, 
                               levels = c(1,2,3,4,5,6), 
                               labels = c("Never",
                                          "Once every three months",
                                          "Once a month",
                                          "Once a week",
                                          "A couple of times per week",
                                          "Every day")),
    avg_reaction_time = as.numeric(avg_reaction_time),
    missed_perc       = as.numeric(missed_perc),
    d_perc_practice   = as.numeric(d_perc_practice),
    d_perc_similarity = as.numeric(d_perc_similarity),
    d_perc_liking     = as.numeric(d_perc_liking)
  )
```

## Adding comparison identifier
```{r}
df_long <- df_long %>%
  dplyr::mutate(
    # Create a consistent order for the stimulus filenames
    stim1 = pmin(left_stimulus, right_stimulus),
    stim2 = pmax(left_stimulus, right_stimulus),
    # Create a pair key that combines the two filenames
    pair_key = dplyr::if_else(
      round_type == "practice",
      paste("practice", stim1, stim2, sep = "_"),
      paste(stim1, stim2, sep = "_")
    )
  ) %>%
  # Create a unique comparison ID by converting the pair key into a factor
  dplyr::mutate(
    comparison_id = as.integer(factor(pair_key))
  )

df_long <- df_long %>% dplyr::select(-stim1, -stim2, -pair_key)
```

Also creating a df for convenience, to lookup with ID corresponds to which comparison
```{r}
df_lookup <- df_long %>%
  select(round_type, left_stimulus, right_stimulus, comparison_order, comparison_id) %>%
  distinct()
```


## Saving to .csv and .Rdata
If you are working with R, you can directly load `load("../data/working/combined_data.RData")`.

```{r}
# Ensure the directory exists
working_dir <- "../data/working"
if (!dir.exists(working_dir)) {
  dir.create(working_dir, recursive = TRUE)
}

# Save trial-level and participant-level data frames to CSV files
write.csv(df_long, file = file.path(working_dir, "df_long.csv"), row.names = FALSE)
write.csv(df_wide, file = file.path(working_dir, "df_wide.csv"), row.names = FALSE)

# Save both data frames in a single .RData file
save(df_long, df_wide, df_lookup, file = file.path(working_dir, "combined_data.RData"))
```

